---
title: "mini-projet - serie temporelle"
author: "XU Yijie, ZHONG Ming (groupe MAM4_SD)"
date: "28/03/2019"
---

objective: estimate the deterministic component of a time series

data source: https://datamarket.com/data/set/22ky/monthly-total-number-of-pigs-slaughtered-in-victoria-jan-1980-august-1995#!ds=22ky&display=line

tools: filters and difference operators

-------------------------------------------------------------

initialization

```{r setup, include=FALSE}
# import time series data
tsData <- read.csv("monthly-total-number-of-pigs-sla.csv")

# remove rows with missing values
tsData <- na.omit(tsData)

# The first column of data is the year and month, ex: '1980-09'
# the second column of data is the number of pigs slaughtered each month. (in Victoria. Jan 1980 – August 1995)
num_pig <- tsData[,2]
# we assume that the period d of the seasonal component is 12
d = 12 
pig_ts <- ts(num_pig,frequency = d)
```

Plot time series

```{r}
ts.plot(tsData,ylab = expression(X[t]))
```


```{r}
decom_ts<- decompose(pig_ts)
plot(decom_ts)
```
```{r}
# the expectation of random noise component E(Y_t)
mean(na.omit(decom_ts$random))
```

E(Y_t) = 98.63634
then we can replace m_t and Y_t with m_t + E(Y_t) and Y_t − E(Y_t) respectively.
 

Estimation and Elimination of Trend in the Absence of Seasonality : 

-------------------------------------------------------------

Trend Estimation


trend estimation 1 : Moving average filter

```{r}
# n : length of data = 188
n = length(pig_ts)
# Let q be a nonnegative integer
q = 2
# add data to the edge
x_t <- c(rep(pig_ts[1], q), pig_ts, rep(pig_ts[n], q))
w_t <- rep(0,n)


for (i in q+1:n+q) {
  for (j in -q:q) {
    w_t[i] = w_t[i] + x_t[i+j]
  }
}
w_t = w_t/(2*q+1)
ts.plot(tsData,ylab = expression(X[t]))
lines(w_t,col="red")
```

trend estimation 2 : Exponential smoothing

```{r}
# fixed α ∈ [0, 1],
alpha = 0.3
# the one-sided moving averages m_t_hat
m_t_hat <- rep(0,n)

# mˆ1 = X1
m_t_hat[1] = pig_ts[1]
m_t_hat[2] = pig_ts[1]
for (i in 3:n) {
  for (j in 0:i-2) {
    m_t_hat[i] = m_t_hat[i] + alpha*(1-alpha)^j*pig_ts[i-j]
  }
  m_t_hat[i] = m_t_hat[i] + (1-alpha)^(i-1)*pig_ts[1]
}

m_t_hat = m_t_hat / 2

ts.plot(tsData,ylab = expression(X[t]))
lines(m_t_hat,col="red")
```

trend estimation 3 : Polynomial fitting (use the method of least squares)

```{r}
# we suppose the trend of the form : m_t = a0 + a1*t + a2*t^2
t = 1:length(num_pig)
tt = t^2
a <- lm(num_pig~t + tt)$coefficients
m_t_poly = a[1] + a[2]*t + a[3]*tt
ts.plot(tsData,ylab = expression(X[t]))
lines(m_t_poly,col="red")
```

It seems that it is under-fitting，so we increase the degree

```{r}
# we suppose the trend of the form : m_t = a0 + a1*t + a2*t^2 + a3*t^3
ttt = t^3
a2 <- lm(num_pig~t + tt + ttt)$coefficients
m_t_poly_2 = a2[1] + a2[2]*t + a2[3]*tt + a2[4]*ttt
ts.plot(tsData,ylab = expression(X[t]))
lines(m_t_poly_2,col="red")
```

-------------------------------------------------------------

Trend Elimination by Differencing

```{r}
# if m_t is a polynomial of degree k
k = 3
pig_diff <- diff(num_pig,differences = k)
plot.ts(pig_diff)
```

-------------------------------------------------------------

Validate with the commands of some time series packages.

```{r}
library(tseries)
library(forecast)
library(itsmr)
# check if our time series data is stationary with Augmented Dickey–Fuller (ADF) t-statistic test
adf.test(pig_ts, alternative="stationary", k=0)
```

Test for moving average filter

```{r}
w_t_package = smooth.ma(pig_ts,q)
ts.plot(tsData,ylab = expression(X[t]))
lines(w_t,col="red")
lines(w_t_package,col="green")
```

Test for exponential smoothing

```{r}
m_t_hat_package <- HoltWinters(num_pig, beta=FALSE, gamma=FALSE)
ts.plot(tsData,ylab = expression(X[t]))
lines(m_t_hat,col="red")
lines(m_t_hat_package$fitted[,1],col="green")
```

-------------------------------------------------------------

Estimation and Elimination of Both Trend and Seasonality : 

-------------------------------------------------------------

the period d of the seasonal component is 12

Estimation of Trend and Seasonal Components

```{r}
# the trend is first estimated by applying
# a moving average filter specially chosen to eliminate the seasonal component and to dampen the noise
q = d/2
n = length(pig_ts) 
x_t <- c(rep(pig_ts[1], q), pig_ts, rep(pig_ts[n], q))
m_t_hat <- rep(0,n)

for (i in q+1:n+q) {
  for (j in -q+1:q-1) {
    m_t_hat[i] = m_t_hat[i] + x_t[i+j]
  }
  m_t_hat[i] = m_t_hat[i] + 0.5*x_t[i+q] + 0.5*x_t[i-q]
}
m_t_hat = m_t_hat/d
# we compute the average w_k of the deviations (x_(k+jd)−m_(k+jd))
w_k = x_t-m_t_hat
# we estimate the seasonal component s_k ：
sum_w_k = 0
for (i in 1:d){
  sum_w_k = sum_w_k + w_k[i]
}
s_k = w_k - d^(-1)*sum_w_k
# the deseasonalized data is then defined to be the original series with the estimated seasonal component removed
d_t = x_t - s_k
# Random = Time series – Trend – Seasonal
y_t_hat = x_t - m_t_hat -s_k
d_t = d_t - mean(na.omit(y_t_hat))
ts.plot(tsData,ylab = expression(X[t]))
lines(d_t,col="red")
```

-------------------------------------------------------------

Elimination of Trend and Seasonal Components by Differencing

```{r}
# if m_t is a polynomial of degree k
k = 3
# the period d of the seasonal component is d
d = 12 

# lag = the period d of the seasonal component = d,  differences = degree of polynomial = k
pig_diff <- diff(diff(num_pig,lag = d,differences = 1),lag = 1,differences = k-1)
plot.ts(pig_diff)
```
